Bird beak's elaboration and exploration

### 2021-06-03 Analysis/03-elaboration_exploration_analyses.Rmd
Updated analyses with within clade proj/rej

## Plotting the ellipses
plot.ellipses(sub_sample_matrices, col = c("grey",colour_vector),
              add = FALSE, transparent.scale = 0.7)
plot.axes(sub_sample_major_axes, col = c("grey",colour_vector),
              add = TRUE, transparent.scale = 0.7)
legend("topleft", legend = c("phylogeny", levels(demo_data$clade)), col = c("grey", colour_vector), lty = 1)
```

@@

```{r, fig.height = 18, fig.width = 18}
## Plotting it
plot.space(demo_data,
           col = colour_vector,
           levels = demo_data$clade,
           xlab = "PC1 (90.5%)",
           ylab = "PC1 (6.86%)",
           xlim = c(-2, 2),
           ylim = c(-1, 1))
points(do.call(rbind,group_centres)[,c(1,2)],
       pch = 13, col = c(colour_vector, "grey"))
plot.axes(major_axes["animal"], col = "grey",
          add = TRUE, transparent.scale = 0.01,
          centre = 0)
plot.axes(major_axes["animal:clade1"], col = colour_vector[1],
          add = TRUE, transparent.scale = 0.01,
          centre = colMeans(demo_data[demo_data$clade == "gulls", c(1,2)]))
plot.axes(major_axes["animal:clade2"], col = colour_vector[2],
          add = TRUE, transparent.scale = 0.01,
          centre = colMeans(demo_data[demo_data$clade == "plovers", c(1,2)]))
plot.axes(major_axes["animal:clade3"], col = colour_vector[3],
          add = TRUE, transparent.scale = 0.01,
          centre = colMeans(demo_data[demo_data$clade == "sandpipers", c(1,2)]))
legend("topleft", legend = c("phylogeny", levels(demo_data$clade)), col = c("grey", colour_vector), lty = c(1,1,1,1), pch = c(NA, 19, 19, 19))
```

@@

In practice we followed [this procedure](https://math.stackexchange.com/questions/598750/finding-the-rotation-matrix-in-n-dimensions) and applied a modification of [this implementation](https://stackoverflow.com/questions/42520301/find-rotation-matrix-of-one-vector-to-another-using-r/42542385#42542385) (see @aguilera2004 for the formal generalisation of this algorithm) using the following algorithm:

@@

```{r, fig.height = 10, fig.heigh = 10}
## Plot the blob wise metrics
plot.analyses.group(group_results, col = colour_vector)
```

@@

```{r}
## Plot results
plot.analyses.tip(tip_results_global, col = colour_vector, group = split(rownames(demo_data), f = demo_data$clade))
```

@@

We also looked at the exploration elaboration profiles within each clades by replacing for each clade, the projection of the elements onto $\vec{B}$ to the projection onto their respective clade major axes $\vec{a}$.

TODO!

>>>>

## Plotting the ellipses
plot.ellipses(sub_sample_matrices, col = c("grey",colour_vector),
              add = FALSE, transparent.scale = 0.5)
plot.axes(sub_sample_major_axes, col = c("grey",colour_vector),
              add = TRUE, transparent.scale = 0.7)
legend("topleft",
       legend = c("phylogeny", levels(demo_data$clade)),
       col = c("grey", colour_vector), lty = 1)
```

@@

```{r, fig.height = 10, fig.width = 10}
## Plotting it
plot.space(demo_data,
           col = colour_vector,
           levels = demo_data$clade,
           xlab = "PC1 (90.5%)",
           ylab = "PC1 (6.86%)",
           xlim = c(-2, 2),
           ylim = c(-1, 1))
points(do.call(rbind,group_centres)[,c(1,2)],
       pch = 13, col = c(colour_vector, "grey"))
plot.axes(major_axes["animal"], col = "grey",
          add = TRUE, transparent.scale = 0.01,
          centre = 0)
plot.axes(major_axes["animal:clade1"], col = colour_vector[1],
          add = TRUE, transparent.scale = 0.01,
          centre = colMeans(demo_data[demo_data$clade == "gulls", c(1,2)]))
plot.axes(major_axes["animal:clade2"], col = colour_vector[2],
          add = TRUE, transparent.scale = 0.01,
          centre = colMeans(demo_data[demo_data$clade == "plovers", c(1,2)]))
plot.axes(major_axes["animal:clade3"], col = colour_vector[3],
          add = TRUE, transparent.scale = 0.01,
          centre = colMeans(demo_data[demo_data$clade == "sandpipers", c(1,2)]))
legend("topleft",
       legend = c("phylogeny", levels(demo_data$clade)),
       col = c("grey", colour_vector),
       lty = c(1,1,1,1),
       pch = c(NA, 19, 19, 19))
```

@@

In practice we followed [this procedure](https://math.stackexchange.com/questions/598750/finding-the-rotation-matrix-in-n-dimensions) and applied a modification of [this implementation](https://stackoverflow.com/questions/42520301/find-rotation-matrix-of-one-vector-to-another-using-r/42542385#42542385) (see @aguilera2004 for the formal generalisation of this algorithm) using the following algorithm implemented in `dispRity::projections` (@disprity):

@@

```{r}
## Plot the blob wise metrics
plot.analyses.group(group_results, col = colour_vector)
```

@@

```{r}
## Plot results
plot.analyses.tip(tip_results_global,
                  col = colour_vector,
                  group = split(rownames(demo_data), f = demo_data$clade),
                  main = "Tips vs. whole phylogeny")
```

@@

Finaly, we also looked at the exploration elaboration profiles within each clades by replacing for each clade, the projection of the elements onto $\vec{B}$ to the projection onto their respective clade major axes $\vec{a}$.

```{r}
## Results within groups
tip_results_within <- analyses.tip(
                    data = demo_data[, c(1:4)],
                    axes = major_axes[-length(major_axes)],
                    group = split(rownames(demo_data), f = demo_data$clade))
```

```{r}
## Plot results
plot.analyses.tip(tip_results_within,
                  col = colour_vector,
                  group = split(rownames(demo_data), f = demo_data$clade),
                  main = "Tips vs. clade")
```
